### 2012-10-18 docs/roadmap/benchmarks.tex
Extracted some benchmarks.

\input{style.sty}

\title{Benchmarks}
\begin{document}
\maketitle
\pagestyle{headings}

The goal of this document is to see how we are comparing with other papers in terms of performance and to note performance progress of the improvements.

\subsection*{Graphic cards}
\def\unt#1{& \footnotesize #1}
\begin{center}\begin{tabular}{lrrr} \toprule
\bf Paper		&				& \bf -- 		& \bf  ATLP\cite{gpu_atlp} \\ \midrule
Serie		&				& GeForce	& Tesla \\
Model		&				& GT 650M	& C1060 \\
Architecture	&				& Kepler		& GT200 \\
Capability		&				& 3.0		& 1.3 \\
Memory \unt{Mb}				& 1024		& 4096 \\
CUDA cores &					& 384		& 240 \\
Core clock \unt{MHz}			& 756		& 1300 \\
Memory clock \unt{MHz}			& 1953		& 1600 \\
Memory bus \unt{bit}				& 128		& 512 \\
Memory bandwidth \unt{GB/s}		& 28.8		& 102.4 \\
Processing power \unt{GFLOPS}	&564.5		& 622.08 \\ \midrule
\bf Processing speedup & 		& 1			& 1.07 \\
\bf Memory speedup & 			& 1			& 3.55 \\ \bottomrule
\end{tabular}\end{center}

\subsection*{Results}
\subsubsection*{ATLP\cite{gpu_atlp}}
\begin{center}\begin{tabular}{lrrrrrrrrrr} \toprule
\bf Matrix size & 128 & 256 & 512 & 1024 & 1536 & 2048 & 2560 & 3072 & 3584 & 4096 \\ \midrule
\bf No split & 0.07 & 0.09 & 0.19 & 0.59 & 1.27 & 2.25 & 3.51 & 5.07 & 6.92 & 9.06 \\
\bf Split at 1 & 0.06 & 0.07 & 0.08 & 0.14 & 0.26 & 0.47 & 0.77 & 1.21 & 1.80 & 2.57 \\ \bottomrule
\end{tabular} \\[4pt]
Matrix multiplication timing in seconds
\end{center}

\bibliographystyle{plain}
\bibliography{bibliography.bib}

\end{document}