### 2012-11-13 docs/roadmap/draft.tex

\documentclass[11pt]{article}
\input{inc/style.sty}

\title{DynaProg for Scala}
\subtitle{A Scala DSL for Dynamic Programming on GPU and FPGA}
\begin{document}
%\maketitle
\shorttitle

\subsection*{Abstract}
Dynamic programming is a common pattern of Computer Science used in various domains. Yet underlying matrix recurrences might be difficult to express and error prone. Additionally, domain experts might not have the skills to make an efficient parallel implementation. In this paper, we leverage the Scala / LMS infrastructure to implement a domain specific language (DSL) for dynamic programming on heterogeneous platforms, which allows to write concise and efficient programs. Our contributions are: \ul
\item Classification of DP problems characteristics (matrix shape, dependency graph, ...)
\item State of the art parallel implementation of these classes on GPUs
\item Grammar embedded in Scala (DSL) to express DP problems efficiently
%\item Normalization of the grammar into efficient productions
\item Systematic approach to generate and process backtracking information
\item Code generator to transform a grammar into efficient code for CPU, GPU and FPGA
\ule

\pagestyle{headings}
\setcounter{tocdepth}{2} \tableofcontents

% ------------------------------------------------------------------------------------------------
\newpage
\section{Introduction}
\subsection{Dynamic programming}
Dynamic programming consists of solving a problem by reusing subproblems solutions. The most famous example of dynamic programming is the Fibonacci series that is defined as 
\[F(n+1) = F(n)+F(n-1) \qquad \text{ with } F(0)=F(1)=1 \]

A typical characteristic of dynamic programming is that a subproblem solution is often reused multiple time, so it is often efficient to store intermediate results in a table to speed-up computation of larger problems.

There exist various categories of dynamic programming:\ul
\item Series, knapsack : that operates usually on a single dimension
\item Sequences alignment (matching two sequences at best), top-down grammar analysis, ...
\item Tree-related algorithms: phylogenetic trees raking, maximum tree independent set, ... 
\ule

Since the first category is inherently sequential and the third category is hard to parallelize efficiently, we focus on the second type of problems, which is also the most common.

{\color{red}
parallel tree-raking, does not share much with other algorithms (sparse version of computations, might not scale efficiently). Most common patterns are already enclosed by the above problems. Real input size is around 300K. We might want to also look at an $O(n^3)$-space-complex problem (like matching 3 strings $S,T,U$).

All the problems we consider use 2D storage matrix, their dependencies are an union of: \ul
\item Serial dependencies
\item Non-serial horizontal or vertical dependencies (1D non-serial)
\item Non-serial horizontal+vertical dependencies in the form $M_{(i,j)} = {\rm op}_k f ( M_{(i,k)}, M_{(k,j)} ) $
\item We have not found other type of dependencies in the literature
\ule}

\subsection{Scala and LMS}

% ------------------------------------------------------------------------------------------------
\input{problems}
\input{design}
\input{implementation}
\input{benchmarks}
\section{Future work}
\section{Conclusion}

% ------------------------------------------------------------------------------------------------
\newpage
\section{Planning/draft}
\subsubsection*{Roadmap}
\begin{tabular}{rl}
16.11 & Rules normalization and automatic backtracking as in 5.2 \\
	& GenScala on LMS + GenCuda + LMS CudaCompiler \\
23.11 & Problem generalization: "cyclic keyword", Zucker problem / CudaLoop optimization \\
30.11 &--- Gap due to LMS missing knowledge \\
7.12 & Benchmarking, grammar analysis \\
14.12 & \\
21.12 & \\
28.12 & --- holiday --- \\
 4.01 & --- holiday --- \\
11.01 & Writing report: implementation description and plan for future work \\
18.01 & Writing report
\end{tabular}

\subsubsection*{Todo @TCK}
\begin{verbatim}
Look at string templating. """xxx $var xxx"""
- Cleanup LMS CUDA generator
- CPU+GPU implementation for serial (NS?) problems larger than device memory

2. separate initialization (base cases, atoms) and processing (rules, recurrences) => less cases to handle (no if)

Plan:
We need to extend the language to support all 3 cases
- Parallelogram is triangular + cyclic => add a cyclic keyword that applies to the whole problem
- Square: what is the transform being done in ADP extension?

1. optimize: terminals of bounded yield + binary decomposition if possible
2. transform into: axioms (init/fixed) + rules (iterations)
3. transform to code
     axiom(i := k(i)) -> M[i,i]=k(i)
     rule(x~y := h(x,y)) -> M[i,j]=forall_i<k<j h(M[i,k],M[k,j]) for both unbounded yield
     rule(x~y_l := h(x,y)) -> M[i,j]=forall_j-l<k<j h(M[i,k],M[k,j]) for y bounded by l
     rule(x~y_l := h(x,y)) -> M[i,j]=h(M[i,j-l],M[j-l,j]) for y of exactly l

XXX: how to encode multi-dimensional matrices
1. assume they have the same type put one after another => different dimensions ok
2. assume of same size => put into a struct
=> but using different pointers seems more reliable => completely different matrices => fixed list of matrices by dimensionality (O(1), O(n), O(n^2), ...) of structs (determined by number of indices to access object)
\end{verbatim}

\subsubsection*{Todo @Manohar}
\begin{verbatim}
- generate chain matrix recurrences to generate implementation
- more theory behind what we want to support, how we encompass all cases, ...
\end{verbatim}

\subsection*{Plan} \ol

\item \textbf{User facing language:} goals are flexibility and compactness.\ul
	\item User-facing language should be similar to related paper \cite{adp_gpu} or \href{http://hackage.haskell.org/package/ADPfusion}{\it ADP fusion}. We want to reuse the transformation mapping (problem description) $\mapsto$ (kernel implementation) for a single element.
	\item We also may want to try to make implicit transformation for code like \\
		{\tt @DP def Fib(n:Int) = if (n<=2) return 1 else Fib(n-1)+Fib(n-2)}.
	\item Windowing: the user should be able to force a windowing (i.e. force a non-serial problem to be a $k$-polyadic serial problem).	
	\item 3 different cases: we care about backtrack cost or both.
	\item Backtracking: create an operator that produces the whole backtrack sequence indices.
	\ule

$\implies$ \emph{end of October}.

\item \textbf{Prototyping:} get a prototype to understand difficulties and share common base. \\ Implement a working prototype of \nameref{aswat} on CPU (for correctness), and specific platform (CUDA/FPGA). This will give us an idea of how to implement the general case. We also need to benchmark and compare both implementations to see how we compare to existing implementations and see the direction to take (which decide is faster and by how much). Here we aim to do as good an implementation for the specific platform (CPU/GPA) as possible.\\
$\implies$ \emph{end of October}.

\item \textbf{Baseline:} Also use benchmarks provided by existing implementations as baselines.\\
$\implies$ \emph{end of October}.

\item \textbf{Formalize IR:} describe the intermediate representation, formalize the framework provided to the code generators (i.e. memory management, ...).
\item \textbf{Full compiler stack:} enrich the compiler stack from both top-down (translate best user-facing language parsers) and bottom-up (parametric code generators), core of the work.
\item \textbf{Benchmark:} make sure implementations are correct, compare them other papers.
\item \textbf{Optimizations:} improve as much as possible / as long as time permits
\ole

\begin{verbatim}
ideas.txt
---------------------------------------------------
Plan:
1. Define the main problem in focus and its class
   - Dimension
   - Serial/non-serial
   - Monadic/polyadic
2. Define some other problems that belong to this class or simpler classes
3. See the feasibility for FPGA (possibly helped by CPU)
4. Implement this algorithm in both CUDA and FPGA
   -> specify all parameters so that we can compare both implementations
5. Find a common IR for the problems
6. Write a code generator for both architectures from IR
   -> make sure all the test-DP are compiled well from IR
7. Find an expressive and compact language for the user-facing representation

Translation into C++:
http://bibiserv.cebitec.uni-bielefeld.de/macports/resources/download/

----------------------------------

Ideas for the language:
- we want maximum flexibility
- backtrack can be custom, we may want just to give back indices
- two problems: we care about cost and we care about whole backtrack.
	@DP def Fib(n:Int) = if (n<=2) return 1 else Fib(n-1)+Fib(n-2)
  should be translated into the more complete syntax that allows more flexibility
  => get inspiration from papers to write an internal language

- we may want to keep track of indices but we do not want to give access to the backtrack matrix / cost matrix

Language side:
- define a simple way to represent some DP programs
- define a comprehensive way to define them => micro-syntax
- define the mapping into IR (actually the IR might BE the micro-syntax)
- define optimizations to IR
  - making serializable by using aggregation functions/transformations
  - avoiding the cost matrix by moving it into the wavefront
- scheduling (with memory loads for non-serializable)
- code generation for both platforms
  - possibly choice to use which platform for what part (load/compute @ cpu/gpu/fpga)

----------------------------------

Core function F:
- in: s,t strings
- in: neighbor costs: top, left, top+left
- in: neighbor stats: top, left, top+left
- out: backtrack information
- out: cost(i,j)
- out: stats

To be most efficient, we need to get exact bit-sizes of
- strings' char
- costs: min/max values
- stats: min/max values
- backtrack

=> we might want to parallelize processing within a thread that operate on 32 or 64 bit inputs (to reduce memory accesses)

Some ideas:
- define a way to pack the characters => less memory transfer (i.e. GATC=>4 letters in 1 char)
- operate on some larger word (ex 64 bits) to increase thread locality and reduce memory accesses
- write a kernel that takes stats from left and r

		stats (x)
		 ||
		 vv
stats -> KK -> new_stats (y')
(y)		 || \
		 vv  --+ backtrack info (Bxy)
		 new_stats (x')

and compute the backtracking for its cell (optional as template boolean)

- multi-grain
  + within the 64b-words : multiple letters at once (thread level)
  + group threads into blocks that operate on (block level)
  + kernel that operates one after another (GPU/CPU level)
    => can we catch a stream's event at CPU level to alloc memory to get back results ?
- keep track of
  + stats O(m+n)
  + backtrack O(m*n)

- are we sure we need squares ?
  - rectangles can increase the full usage of all threads during (x-y) runs
  - assuming we put longest word vertically we play hot potato for stats
    left->right => block.x exchanges within block, last thread writes to
    global exchange memory (if short, so that it does not penalize running time)
  - we may go up to running in stripes
- if we put in diagonal-major data in the memory, why not stream all chunks like
  that using a cyclic buffer (?) => no empty slot
  ==> this might be useful for the backtrack information

- use extensively profiling : CUDA profiler

- Robust DP paper provides 2 insights: coalesced access + GPU synchronization
  => can we do a producer-consumer scenario at the block border so that we can
     execute a large diagonal horizontal/vertical swipe made of multiple blocks?
  => multiple swipes at different delays (so that we avoid syncing at every step

          s-->
  +----------------/------------+
t | B1            /             |
| +------/-------/              | KERNEL1 (keep benefit of t in shared mem)
| | B2  /                       | delay between b1 and b2 can be 1% of line length
v +----/------------------------+
  |                             |
  |                             | KERNEL2
Memory organized as
b1 -- --
b1 b1 --
b1 b1 b1
b1 b1 b1
.. .. ..
b1 b1 b1
b2 b1 b1
b2 b2 b1
b2 b2 b2
.. .. ..

For stats (aggregation of the table): we need to maintain both an horizontal and vertical front
For backtracking we need to maintain the whole table
     .. | . |
	..A | B | min/max: Min(D)=Min(Min(C),Min(B))
	----+---+ sum    : Sum(D)=Sum(C)+Sum(B)-Sum(A)
	..C | D | avg    : sum both #cells and values then divide at appropriate cell
	----+---+
we many not need to store what's the previous cell, however, the previous
cell information (usually limited range) is much more compact than the score(32-64 bits)

Problem parameters:
- bit size (alphabet,stats,cost,backtrack)
- dimensionality (2D,3D for the matrix)
  - 3D problem example is Smith Waterman with arbitrary gap weight function.
- Summarizability of the set: for computing the stats of an element, we can limit ourselves to computing values up to a certain value
  -in SW with gap extending penalty, we can summarize easily.
  -in Nussinov, we _need_ to access all the previous row-column elements for the pairwise sum (without a window limit).

user-interface.txt
--------------------------------------
Reuse most of the syntactic elements and write it as uneducated programmer, apply annotations/additions on top of it or detect that it is DP and then converge to our system.

Smith-Waterman (from problems of interest)
--------------------------------------------

trait SWOps {
	type T
	type Cost = Int
	type Backtrack = (i,j) => Direction
	def rightGap(_ :_, c:Char) : [Cost, Backtrack]
}

def myRec = {
	def M[i,j] : Int = max {
		M[i,j-1] + gapRight,
		M[i-1, j] + gapLeft,
		M[i-1,j-1] + delta(S(i),T(j))
	}
}

def SWat(S0: List[Char], T0:List[Char]) : List[Pair[Char,Char]] = {
	def cost(s: Char, t: Char) = if (s==t) 2 else 1 // example cost function

	def DP(S: List[Char], T:List[Char]): [Score,Backtrack,Output[Pair[Char,Char]] = (S,T) match {
		case (_, Nil) => (0, stop)
		case (Nil, _) => (0, stop)
		case Si::Sr, Tj::Tr) =>
			val nw = SWat(Sr,Tr);
			val n  = SWat(Sr,T);
			val w  = SWat(S,Tr);
			val c = cost(Si,Tj);

			if (nw._1+c>=max(n,w)-d) then (nw._1+c,"NW",[Si,Tj])
			else if (n>=w) then (n._1-d, "N",[Si,_])
			else (w._1-d, "W",[_,Tj])
	}
	backtrack(DP(S0,T0))
}

- Char, Score, Backtrack, Output are LMS Rep objects so that we can easily detect
  when we need to use dynamic programming.
- backtrack() is a new function that take the Backtrack information and reconstruct
  the desired output backwards

We would unfold this program into two:
1. The dynamic programming that constructs all the solution and that is the core of the algorithm
2. The backtrack function that reconstruct the solution and can be run anywhere (Java, CPU, ...)

URLs.txt
------------------------------------
Some GPU algorithms:
http://hgpu.org/?cat=11

CUDPP libraries (but awfully big resulting binary):
http://code.google.com/p/cudpp/

13 dwarfs:
http://developer.amd.com/afds/assets/presentations/2155_final.pdf

http://tutorials.jenkov.com/java-reflection/fields.html
http://lampwww.epfl.ch/~michelou/scala/scala-reflection.html

XXX: use TypeClass to put a predicate on types
--------------------------------------------
def fun[T: CanTranslateToC](...)
def fun[T](implicit ev:CanTranslateToC[T])

class CanTranslateToC[T] { def translate:String }
implicit def canTranslateInt = new CanTranslateToC[Int] = { def translate = "Int" }
\end{verbatim}

% ------------------------------------------------------------------------------------------------
\newpage
\bibliographystyle{finplain}
\bibliography{bibliography.bib}
\end{document}