### 2012-11-13 docs/roadmap/implementation.tex

\newpage
\section{Implementation}
% ------------------------------------------------------------------------------------------------
\subsection{ADP parsers}
\subsection{Transformations and simplifications}
\subsection{LMS integration}
\subsection{CUDA implementation}