### 2012-11-28 docs/roadmap/implementation.tex
Nits.

\newpage
\section{Implementation}
% ------------------------------------------------------------------------------------------------
\subsection{ADP parsers}
\subsection{Transformations and simplifications}

\begin{verbatim}
 * Parser construction:
 * 1. Assign a "OR_id" (sub_rule_id) and "CONCAT_id" to all parsers so that we know which rule applies
 *    How to skip some indices ?
 *
 * Scala running:
 * 2. Provide meaningful backtrack: rule_id and list of concat indices
 *
 * Code generation:
 * 1. Normalize rules (at hash-map insertion (?))
 * 2. Compute dependency analysis between rules => order them into a list/queue
 *    - If rule R contains another rule S unconcatenated (or concatenated with empty)
 *      then we have S -> T (S before T)
 * 3. Compute the maximal number of concatenation among each rule (field in Treeable(?))
 * 4. Break rules into subrules (at each Or, which must be at top of the rule)
\end{verbatim}

\subsection{LMS integration}
\subsection{CUDA implementation}