### 2013-01-17 docs/report/report.tex

\documentclass[11pt]{article}
\input{inc/style.sty}
\usepackage{multicol,etoolbox}

\title{DynaProg for Scala}
\subtitle{A Scala DSL for Dynamic Programming on CPU and GPU}
\begin{document}
\maketitle
\input{introduction}
\input{problems}
\input{design}
\input{implementation}
\input{benchmarks}
\input{conclusion}

% ------------------------------------------------------------------------------------------------
%\section*{Planning}
%Legacy roadmap deadlines: \\
%\begin{tabular}{ll}
%Nov 16 & Rules normalization and automatic backtracking \\
%	& GenScala on LMS + GenCuda + LMS CudaCompiler \\
%Nov 23 & Problem generalization: "cyclic keyword", Zucker problem / CudaLoop optimization \\
%Nov 30 &--- gap due to LMS missing knowledge --- \\
%Dec 7 & Benchmarking, grammar analysis \\
%Dec 14 & First thoughts for larger than device memory \\
%Dec 21 & Writing report \\
%Jan 4 & --- holiday --- \\
%Jan 18 & Writing report: implementation description and plan for future work
%\end{tabular}
%
%\subsubsection*{Journal}
%\begin{tabular}{ll}
%Sep. 17	& Getting started with the project, reading related work on hash maps. \\
%Sep. 24	& Parallel hashing paper solve the problem for 32bit key/value pair. Stripped CudPP. \\
%		& Devised (but not implemented) extension beyond 64bit using memory areas locks. \\
%---		& Change of project suggested by Vojin (supervisor): joint work with Manohar on DP \\
%Oct. 01	& Problems specifications: serial/non-serial, started CUDA implementation with blocks. \\
%Oct. 08	& CudAlign solves serial monadic, might adapt it (but complicated / ad-hoc / in progress). \\
%		& Focus on non-serial problems that fit in GPU memory. \\
%Oct. 15	& First working implementation for non-serial problems (rectangle, triangle, parallelogram) \\
%Oct. 22	& No workaround for timeout, fixed by multiple kernels. Implemented backtrack on GPU. \\
%Oct. 29	& Scala/C compiler engine, to use Scala/CUDA, C code must be provided. ADP (Manohar) \\
%Nov. 05	& Multiple fixes and rework of ADP parsers to aim at generating C-like code. \\
%Nov. 12	& Rework ADP parsers: cyclic, two-track grammars and automatic backtracking discussion. \\
%Nov. 19	& Explorations LMS and Macros for C code, implementation is quite different, hence ad-hoc. \\
%Nov. 26	& Full backtracking stack: apply / unapply / reapply, refactoring of the classes. \\
%Dec. 03	& JNI for LibRNA to get coefficients for Zuker, errors in algebra (quite hairy code). \\
%Dec. 10	& Rework concatenation operators, yield analysis, code generation. \\
%Dec. 17	& Detupling, generic backtrack (vs. ad-hoc), nested aggregates, empty results support. \\
%Dec. 24	& (sick 4 days), code generation: full JNI conversion support, rework CodeCompiler. \\
%Dec. 31	& Writing report (design, implementation), fixing various issues, complexity analysis. \\
%Jan. 7	& Writing report, attempt to fix Zuker, fix SBT, strip LibRNA. LMS C codegen (Manohar) \\
%Jan. 14	& Report improvements, fixed Zuker (coherent with GAPC), {\color{red} XXX} \\
%\end{tabular}

%Some GPU algorithms: http://hgpu.org/?cat=11
%Translation into C++: http://bibiserv.cebitec.uni-bielefeld.de/macports/resources/download/
%CUDPP libraries (but awfully big resulting binary): http://code.google.com/p/cudpp/
%13 dwarfs: http://developer.amd.com/afds/assets/presentations/2155_final.pdf
%http://tutorials.jenkov.com/java-reflection/fields.html
%http://lampwww.epfl.ch/~michelou/scala/scala-reflection.html
%
%Hint: use TypeClass to put a predicate on types
%  def fun[T: CanTranslateToC](...)
%  def fun[T](implicit ev:CanTranslateToC[T])
%  class CanTranslateToC[T] { def translate:String }
%  implicit def canTranslateInt = new CanTranslateToC[Int] = { def translate = "Int" }

% ------------------------------------------------------------------------------------------------
\newpage
\patchcmd{\thebibliography}{\section*{\refname}}{\begin{multicols}{2}[\section*{\refname}]}{}{}
\patchcmd{\endthebibliography}{\endlist}{\endlist\end{multicols}}{}{}
\bibliographystyle{plain} % acm
{\small \bibliography{bibliography.bib}}
\end{document}