### 2013-01-29 docs/slides/res/trans.tex
Up to 2x faster than RNAfold-RL (1.77x on average) with optimization in LibRNA.

\documentclass[11pt]{article}
\usepackage{amssymb,amsmath,amsthm,hyperref,verbatim,pict2e,graphicx,marvosym,array,booktabs}
\usepackage[usenames]{xcolor} % color names ,dvipsnames,svgnames,table
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\pagestyle{empty} % remove page numbers
\oddsidemargin -2.5cm
\evensidemargin -2.5cm
\topmargin -4cm
\headsep 1cm
\parskip 0cm
\parindent 0cm
\setlength{\paperwidth}{14cm}
\setlength{\paperheight}{6cm}
\begin{document}

%\color{white}
\[T < \textbf{Aggregate} < \textbf{Or} < \textbf{Filter} < \textbf{Map} < \textbf{Concat} < \text{(Tabulation | Terminal)}\]

\begin{center}
\begin{tabular}{l|lllll}
Outer $\diagdown$ inner	& Aggregate	& Or			& Map	& Filter		& Concat \\ \hline
Aggregate			& merge?		& swap$_R$	& ---		& swap$_P$	& ---\\
Or					& ---			& simplify?	& ---		& swap$_P$?	& ---\\
Map					& ---			& swap$_R$	& fuse	& swap$_P$	& ---\\
Filter					& ---			& ---			& ---		& merge?		& ---\\
Concat				& ---			& swap$_R$	& ---		& ---			& ---\\
\end{tabular} \\[10pt]
\footnotesize{$R=\ $required, $P=\ $performance optimization, $?=\ $if possible}
\end{center}

\end{document}