### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Meta/bnfMeta.tex

<metarule> ::= <rule name> "::=" <abstract node> ":>" <semantics>
                       | <rule name> "::=" <switch_rule> ":>" <semantics>
<rule name> := <MMAsymbol>
<abstract node> ::= <MMAsymbol> "[" <list of arguments> "]"
<switch_rule> ::= <MMAsymbol> ":>" <list of switch arguments>
<semantics> ::= a call to a semantic function (*)
<argument> ::= <symbol> ":" <pattern>
                        | <MMAsymbol> ":" <rule name>
                        | <MMAsymbol> ":" "{" <rule name> "}"
<switch argument> ::= "_" <MMAsymbol> "->" <rule name>
						| <MMAsymbol> "->" "{" <rule name> "}"
<pattern> ::= any Mathematica pattern