### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/Domlib.tex

%
% File created on {2009, 4, 25, 17, 47, 46.071200} by makeDoc.
%
% Section header
\section{The Alpha`Domlib` package } 
\label{label:Alpha`Domlib`}
\alphanote{Domlib}{Documentation revised on August 10, 2004}

\alphausage{DomLib}{Domlib is a library of domain functions. It contains the functions const2al,
const2mma, dom2mma, dom2al, domCompRays, DomAddRays, DomBasis, DomConstraints,
DomConvex, DomCost, DomDifference, DomEmpty, DomEmptyQ, DomEqualities,
DomEqualQ, DomExtend, DomImage, DomIntersection, DomIntEmptyQ, DomLeftHermite,
DomLTQ, DomMatrixSimplify, DomPreimage, DomProject, DomRays, DomRightHermite,
DomSimplify, DomSort, DomUnion, DomUniverse, DomUniverseQ, DomVertices,
DomVisual, DomZImage, DomZPreimage, hypercube, LatticeDifference,
LatticeImage, LatticeIntersection, LatticeHermite, LatticePreimage,
linearConstraintQ, linearExpQ, linHalfSpace, polToZpol, rays, vertices,
and zpolToPol.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomLib}

\alphanote{Domlib}{Domlib is based upon the PolyLib library, which is a set of public domain 
C programs first developed at Irisa. Domlib is interfaced to this library
using MathLink.}

\alphausage{domlib}{Math link. Link to the external program "domlib". The mechanism of 
MathLink is pretty complicated and needs to be clarified here.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{domlib}

\alphausage{const2al}{const2al[ind,c] translates in Alpha form the constraint c. 
ind is an index list (e.g. \{"i","j"\}) and c is a constraint in
form eq[i+2j,3] or ge[i+2j,3].}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{const2al}

\alphausage{const2mma}{const2mma[indexList,vectorList] converts a constraint, expressed as a list of 
index names and a list of vectors, into Mathematica form}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{const2mma}

\alphausage{dom2mma}{dom2mma[d] converts Alpha domain d into a pair \{constraints,index\} (e.g.
\{\{i,j\},\{i+j$>$=2,...\}\}) suitable for Mathematica.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{dom2mma}

\alphausage{dom2al}{dom2al[\{cst,ind\}] translates a domain given in MMA
form into its Alpha encoding. ind is a list of indexes, and cst is
a list of constraints of the form i+j $>$ 2. Warning: left-hand side is
the linear part, and right-hand side is an integer.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{dom2al}

\alphanote{dom2al}{In dom2al, the symbols must be strings, otherwise there is a problem.
This has to be checked.}

\alphausage{domCompRays}{domCompRays[dom] recomputes the rays of the Alpha domain dom and
returns an Alpha domain. domCompRays ignores the ray part of dom, and
recomputes it. This function allows one to modify a contraint in a domain,
and to update the domain accordingly.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{domCompRays}

\alphausage{domHalfSpaceQ}{domHalfSpaceQ[dom] is True if the Alpha domain dom is a half-space. dom 
is either a string or an ast.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{domHalfSpaceQ}

\alphausage{DomAddRays}{DomAddRays[dom,m] returns the domain dom augmented with rays from the matrix
m. rays are in Alpha format (i.e. n+2 components for n dimensions).

Example:

dom is \{i,j $|$ i $>$= 0\},

m=matrix[3, 4, \{\}, \{\{1, 0, 1, 0\}, \{1, 10, 10, 1\}\}],

DomAddRays[dom,m] returns \{i,j $|$ 0$<$=i$<$=(j,10)\}.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomAddRays}

\alphausage{DomBasis}{DomBasis[m] returns a row basis of the Alpha matrix m using Gauss 
Jordan-elimination.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomBasis}

\alphausage{DomConstraints}{DomConstraints[m] returns the minimum convex polyehdron defined by 
the constraint matrix m. DomConstraints[m1,m2] returns the Z-polyhedron 
obtained by image of the polyhedrom P by the invertible mapping m1, 
where P is the minimum convex Polyhedron satisfying the constraints 
given in m2, i.e., Z = m1(DomConstraints[m2]).}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomConstraints}

\alphausage{DomConvex}{DomConvex[d] returns the minimum convex polyhedron which encloses the
(polyhedral or Z) domain d. Warning (4/11/98), this function was bugged 
in Polylib, and the current implementation uses DomRays instead of 
DomConvex.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomConvex}

\alphausage{DomCost}{DomCost[d, c] returns the interval of values of the cost function c
evaluated over domain d = \{MinN, MinD, MinI, MaxN, MaxD, MaxI\}.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomCost}

\alphanote{DomCost}{DomCost is a function of the Domlib library, and what it does in 
not clear.}

\alphausage{DomDifference}{DomDifference[d1, d2] returns the domain difference of domain d1 less d2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomDifference}

\alphausage{DomEmpty}{DomEmpty[n] returns the empty domain of dimension n.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomEmpty}

\alphausage{DomEmptyQ}{DomEmptyQ[d] returns True if domain d is empty, False otherwise. The
test of emptyness is based uniquely on the rational polyhedron, not the 
integral polyhedron.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomEmptyQ}

\alphausage{DomEqualities}{DomEqualities[d] returns the matrix of equations (lineality space) of
domain d (does not handle union of convexes).}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomEqualities}

\alphausage{DomEqualQ}{DomEqualQ[d] returns True if domain d1 is equivalant to d2, False
otherwise.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomEqualQ}

\alphausage{DomExtend}{DomExtend[dom, idx] extends  dom to the indices in the index list idx. 
This list should contains the indices of dom.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomExtend}

\alphausage{DomImage}{DomImage[d,m] returns the image of the domain d under the transformation 
matrix m. This function always returns a polyhedral domain. 
If d is a Z-Domain, it returns the image of the rational polyhedral 
domain enclosing the Z-Domain.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomImage}

\alphausage{DomIntersection}{DomIntersection[d1, d2] returns the domain intersection of domains d1 and d2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomIntersection}

\alphausage{DomIntEmptyQ}{DomIntEmptyQ[dom1, dom2] returns True if dom1 contains no integral points
in the context of dom2, False otherwise. ??}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomIntEmptyQ}

\alphausage{DomLeftHermite}{DomLeftHermite[m] returns \{H, Q\} where m = HQ, Q unimodular, H hermite.
(Warning, bugged function please use hermiteL[]).}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomLeftHermite}

\alphausage{DomLTQ}{DomLTQ[dom1, dom2, idx, pdim] compares dom1 and dom2 at index position
idx (integer). pdim is the dimension of the parameter space. Returns 1 if
dom1$>$dom2, returns -1 if dom1$<$dom2, returns 0 if dom1$>$$<$dom2 (whatever it
means.)}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomLTQ}

\alphanote{DomLTQ}{DomLTQ does not check the type and number of parameters.}

\alphausage{DomMatrixSimplify}{DomMatrixSimplify[m1,m2] returns m1 simplified in the presence of m2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomMatrixSimplify}

\alphausage{DomPreimage}{DomPreimage[d,m] returns the preimage of the domain d under the transformation 
matrix m. This function always returns a polyhedral domain. For a
ZDomain, it returns the pre-image of the rational polyhedral domain 
enclosing the ZDomain.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomPreimage}

\alphausage{DomProject}{DomProject[dom, idx] projects dom onto the indices given 
in the index list idx. Reorders indices and changes the dimension
accordingly.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomProject}

\alphausage{DomRays}{DomRays[m] returns the minimum convex polyhedron defined by the rays 
given in the matrix m.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomRays}

\alphausage{DomRightHermite}{This function is bugged, use hermiteR instead.
DomRightHermite[m] returns the Hermite decomposition \{Q, H\} of the
Alpha matrix m, i.e. m = QH, Q unimodular, H Hermite.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomRightHermite}

\alphausage{DomSimplify}{DomSimplify[d1, d2] returns a domain equal to d1 simplified in 
the context of d2. In other words, we remove of the definition of
d1 all constraints which are implied by d2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomSimplify}

\alphausage{DomSort}{DomSort[ldom, idx, pdim, time:True$|$False] returns the 
topological ordered list of domains ldom. idx is the level to 
consider for sorting, pdim is the parameter space dimension. 
Returns a list of logical times (one per domain) if time is True, 
otherwise it returns a permutation of ldom. An application of this 
permutation to ldom returns a sorted list.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomSort}

\alphanote{DomSort}{This function seems to be fragile. The only place where it is used 
is in the INorm package.}

\alphausage{DomUnion}{DomUnion[d1, d2] returns the domain union of domains d1 and d2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomUnion}

\alphausage{DomUniverse}{DomUniverse[n]	returns the universe domain of dimension n.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomUniverse}

\alphausage{DomUniverseQ}{DomUniverseQ[d] returns True if domain d is the universe, False otherwise.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomUniverseQ}

\alphausage{DomVertices}{DomVertices[ldom, context] finds the parametrized vertices of a list 
of parametrized polyhedra ldom. context is the domain of 
parameters. It returns a list of pairs whose elements contain 
a parameter domain, and a list of parametrized vertices. 
DomVertices[pol, param] finds the parametrized vertices 
of a parametrized polyhedron pol.
DomVertices[pol], finds the vertices of the non-parametrized 
polyhedron pol. It returns \{\{e,l\}\} where e is the universe 0-domain and 
l is a list of non-parametrized vertices.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomVertices}

\alphanote{DomVertices}{This function seems to be fragile. It is used nowhere in \MMAlpha{}.}

\alphausage{DomVisual}{DomVisual[d1,d2] visualizes the domain d1 with the Opera tool. d2 is
the parameter domain. Warning if the domain d2 is ommited, it is
replaced by the parameter domain of \$result. WARNING: CURRENTLY NOT
AVAILABLE (the function does nothing).}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomVisual}

\alphausage{DomZImage}{DomZImage[dom, mat] finds the Z-image of the domain dom by the matrix mat.
This function may or may not return a Z-Domain depending on the 
transformation matrix.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomZImage}

\alphausage{DomZPreimage}{DomZPreimage[dom, mat] finds the Z-Preimage of the domain dom by the 
matrix mat. This function may or may not return a Z-Domain depending on 
the transformation matrix.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomZPreimage}

\alphausage{hypercube}{hypercube[n] creates a hypercubic domain (in Mathematica form) of dimension n, 
and size 10. dom2l[hypercube[n]] allows such a domain to 
be translated into Alpha form. Useful for testing purposes.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{hypercube}

\alphausage{LatticeDifference}{LatticeDifference[m1,m2] returns the difference of the lattices m1 and m2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{LatticeDifference}

\alphanote{LatticeDifference}{Seems to return a list of matrices.}

\alphausage{LatticeImage}{LatticeImage[m1,m2] returns the image of the lattice m1 by the function m2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{LatticeImage}

\alphausage{LatticeIntersection}{LatticeIntersection[m1,m2] returns the intersection of the lattices m1 and m2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{LatticeIntersection}

\alphausage{LatticeHermite}{LatticeHermite[m] returns the lattice m in Hermite Normal Form.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{LatticeHermite}

\alphausage{LatticePreimage}{LatticePreimage[m1,m2]	returns the preimage of the lattice m1 by the 
function m2.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{LatticePreimage}

\alphausage{linearConstraintQ}{linearConstraintQ[c,\{symbols\}] is True if c is a linear constraint
formed with symbols, False otherwise.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{linearConstraintQ}

\alphausage{linearExpQ}{linearExpQ[exp,\{symbols\}] is True if exp is a linear expression 
formed with symbols, False otherwise.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{linearExpQ}

\alphausage{linHalfSpace}{linHalfSpace[h] returns the Mathematica Matrix corresponding to the Alpha
half-space h.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{linHalfSpace}

\alphausage{polToZpol}{If the domain d is a union of polyhedra, polToZpol[d] returns the
equivalent Z-polyhedron, otherwise it returns d as is.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{polToZpol}

\alphausage{rays}{rays[d] returns the list of rays of the Alpha domain d. rays[\{const,index\}]
returns the vertices of d given in Mathematica form.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{rays}

\alphausage{vertices}{vertices[d] returns the list of vertices of the Alpha domain d. 
vertices[\{const,index\}] returns the vertices of d given in Mathematica form.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{vertices}

\alphausage{zpolToPol}{If the domain d is a Z-Domain, zpolToPol[d] returns the rational polyhedral
domain enclosing d, otherwise it returns d as is.}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{zpolToPol}

\alphausage{zpolIsPolQ}{zpolIsPolQ[d] is True if the Z-polyhedron d is actually a polyhedron
(i.e. has identity matrices as lattices).}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{zpolIsPolQ}

\alphausage{DomTrueRays}{DomTrueRays[ dom ] returns the list of true rays of dom, i.e. rays which
are not lines}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomTrueRays}

\alphausage{DomConstraintsOfDom}{DomConstrainsOfDom[ dom ] returns the constraints }{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomConstraintsOfDom}

\alphausage{DomLines}{DomLines[ dom ] returns the lines of dom}{Alpha/Domlib.m}{Alpha`Domlib`}
\index{DomLines}