### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/INorm.tex

%
% File created on {2007, 11, 15, 17, 10, 19.106120} by makeDoc.
%
% Section header
\section{The Alpha`INorm` package } 
\label{label:Alpha`INorm`}
\alphausage{iNorm}{iNorm[s:\_system, k:\_Integer, options] computes the  imperative form of system s. k is the dimension of time. As usual,  iNorm[k, options] means \$result = iNorm[\$result, k, options]. See  Options[iNorm] for options.}{Alpha/INorm.m}{Alpha`INorm`}
\index{iNorm}

\alphausage{rnf}{rnf[s:\_system] returns the R normal form of s.}{Alpha/INorm.m}{Alpha`INorm`}
\index{rnf}