### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/Lexicographic.tex

%
% File created on {2007, 11, 15, 17, 10, 21.043002} by makeDoc.
%
% Section header
\section{The Alpha`Lexicographic` package } 
\label{label:Alpha`Lexicographic`}
\alphausage{domLexGreater}{domLexGreater[v:\_Matrix] returns a domain D such that for all z in D, z$>$v, where $>$ means lexicographically greater. V is a parametrized vertex. If v is a n*(p+1) matrix, D is a n+p domain, and the p last dimensions are parameter dimensions. Example: domLexGreater[\{\{1,0\},\{0,2\}\}] returns \{a,b,c $|$ c+1$<$=a\} $|$ \{a,b,c $|$ a=c; 3$<$=b\}, where c is the parameter index. domLexGreater[v:\_Vector] is a shortcut for 0-dimensional parameter space. (i.e. domLexGreater[\{0,0,1\}] means domLexGreater[\{\{0\},\{0\},\{1\}\}].}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{domLexGreater}

\alphausage{domLexLower}{domLexLower[v:\_Matrix] returns a domain D such that for all z in D, z$<$v, where $<$ means lexicographically lower. V is a parametrized vertex. If v is a n*(p+1) matrix, D is a n+p domain, and the p last dimensions are parameter dimensions. Example: domLexLower[\{\{1,0\},\{0,2\}\}] returns \{a,b,c $|$ a$<$=c-1\} $|$ \{a,b,c $|$ a=c; b$<$=1\}, where c is the parameter index. domLexLower[v:\_Vector] is a shortcut for 0-dimensional parameter space. (i.e. domLexLower[\{0,0,1\}] means domLexGreater[\{\{0\},\{0\},\{1\}\}].}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{domLexLower}

\alphausage{lexGreaterQ}{lexGreaterQ[a:\_Matrix, b:\_Matrix, p:\_domain] returns True  if a is lexicographicallygreater than b for any value of parameters in p. a and b  are parametrized vectors.}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{lexGreaterQ}

\alphausage{lexGreaterOrEqualQ}{lexGreaterOrEqualQ[a:\_Matrix, b:\_Matrix, p:\_domain] returns True if a is lexicographically greater or equal than b for any value of parameters in p. a and b are parametrized vectors.}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{lexGreaterOrEqualQ}

\alphausage{lexLowerQ}{lexLowerQ[a:\_Matrix, b:\_Matrix, p:\_domain] returns True  if a is lexicographically lower than b for any value of parameters in p. a and b  are parametrized vectors.}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{lexLowerQ}

\alphausage{lexLowerOrEqualQ}{lexLowerOrEqualQ[a:\_Matrix, b:\_Matrix, p:\_domain] returns True if a is lexicographically lower or equal than b for any value of parameters in p. a and b are parametrized vectors.}{Alpha/Lexicographic.m}{Alpha`Lexicographic`}
\index{lexLowerOrEqualQ}