### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/Properties.tex

%
% File created on {2007, 11, 15, 17, 10, 20.514177} by makeDoc.
%
% Section header
\section{The Alpha`Properties` package } 
\label{label:Alpha`Properties`}
\alphanote{Properties}{Documentation checked on August 8, 2004}

\alphausage{Properties}{The Alpha`Properties` package contains the definition of various  property-checking functions. Should be used as the preferred place  for adding new property-checking predicates that might be of large  interest.}{Alpha/Properties.m}{Alpha`Properties`}
\index{Properties}

\alphanote{Properties}{This package should probably merged with other packages such as Tables, Substitution, etc.}

\alphausage{allDomEqualQ}{allDomEqualQ[domList] is True is all the domains in the list domList are equal.}{Alpha/Properties.m}{Alpha`Properties`}
\index{allDomEqualQ}

\alphausage{allDomDisjointQ}{allDomDisjointQ[domList] is True if all the domains in a list domList are two-by-two disjoint.}{Alpha/Properties.m}{Alpha`Properties`}
\index{allDomDisjointQ}

\alphausage{allDomUnion}{allDomUnion[domList] returns the domain which is the union of the domains in domList.}{Alpha/Properties.m}{Alpha`Properties`}
\index{allDomUnion}

\alphausage{uniformQ}{uniformQ[sys] is True if the Alpha program sys is uniform  (all dependences between local variables are translations),  False otherwise. The default value of sys is \$result.  uniformQ can also be applied to an Alpha matrix or to  an affine AST.}{Alpha/Properties.m}{Alpha`Properties`}
\index{uniformQ}