### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/VhdlTestBench.tex

%
% File created on {2007, 11, 15, 17, 10, 20.994002} by makeDoc.
%
% Section header
\section{The Alpha`VhdlTestBench` package } 
\label{label:Alpha`VhdlTestBench`}
\alphausage{VhdlTestBanch}{Package,contains definition of funtion for generating test benches for VHDL generated from Alphard}{Alpha/VhdlTestBench.m}{Alpha`VhdlTestBench`}
\index{VhdlTestBanch}

\alphausage{vhdlTestBenchGen}{vhdlTestBenchGen[sys] generates a VHDL test bench for  the VHDL component generated from sys with the a2v command. sys mustbe a valid Alphard system, it works if sys is a cell, a module or a controller (not an interface), as for a2v, the parameters must be assigned. vhdlTestBenchGen[] applies on \$result}{Alpha/VhdlTestBench.m}{Alpha`VhdlTestBench`}
\index{vhdlTestBenchGen}

\alphausage{mkVhdlLoop}{test}{Alpha/VhdlTestBench.m}{Alpha`VhdlTestBench`}
\index{mkVhdlLoop}

\alphausage{mkVhdlLowerBound}{test}{Alpha/VhdlTestBench.m}{Alpha`VhdlTestBench`}
\index{mkVhdlLowerBound}

\alphausage{mkVhdlUpperBound}{test}{Alpha/VhdlTestBench.m}{Alpha`VhdlTestBench`}
\index{mkVhdlUpperBound}