### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/incl-Alpha.tex

% 
\subsubsection{Introduction}
The \texttt{Alpha.m} package is the main package of Mathematica.
It is located in the directory \texttt{\$MMALPHA/lib/Packages}.
It contains the definition of the main symbols and data structures
of \MMAlpha{}.

A special word about the \texttt{exportAlphaFunctions} function.
This function takes as a parameter the autoloadFile. This file
is located in 
\begin{verbatim}
$MMAlpha/lib/Packages/Alpha/autoload.m
\end{verbatim}
%$
and is created automatically by the \texttt{exportAlphaFunctions}.

At the end of the \texttt{Alpha.m}, a few instructions are executed
when loading this package. If the \texttt{autoloadFile} does not
exists, \texttt{exportAlphaFunctions} is called and creates it.

The function \texttt{exportAlphaFunctions} uses a local variable
\texttt{contexts} of \texttt{Alpha.m} which contains the list of 
the packages that need to be loaded when using \MMAlpha{}. This variable
has to be changed when a new package is added. For each package
of \texttt{contexts}, \texttt{exportAlphaFunctions} creates an 
entry in the autoload file that declares the corresponding package
using the \texttt{DeclarePackage} function of Mathematica. In this
way, all packages are declared, and when a symbol of the package
is evaluated for the first time, 
the corresponding package is automatically loaded.

The remaining of this chapter describes the functions of the 
Alpha package.