### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/ReferenceManual/referenceManual.tex

\documentclass[11pt]{report}
\usepackage{verbatim,version,epic,eepic}
\usepackage{makeidx}
\makeindex

\newcommand{\MMAlpha}{{\sc mmalpha}}
\newcommand{\MMALPHA}{{\sc mmalpha}}

\newcommand{\MMAlfa}{{\sc mmalpha}}
\newcommand{\Alfa}{{\sc alpha}}
\newcommand{\mma}{{\sc Mathematica}}
\newcommand{\alfa}{{\sc alpha}}
\newcommand{\myquotes}{"} 
\newcommand{\myitem}{\\}

\newcommand{\alphausage} [4]
{\subsection{#1} \noindent
{\rm #2}\\{\bf Defined in file:} #3\\{\bf Package:} #4}

\newcommand{\alphanote} [2]
{~\\\noindent{\footnotesize #2}}

\newcommand{\packageusage} [3]
{\subsection{#1} 
{\rm #2}}

\begin{document} 
\pagestyle{headings}
\thispagestyle{empty}
\title{MMAlpha Version 2.0 -- Reference Manual\thanks{Last update of
\today{}}} 
\author{Api/Cosi/R2D2/Compsys/Cairn}

\date{March 2008}

\maketitle

\tableofcontents

\newpage 
\chapter{Introduction}
This document is the reference Manual of \MMAlpha{}. It concerns
Version 2.0 of \MMAlpha{}, but is still under construction.
It has been
(almost) automatically produced from the usage and note statements
of the \MMAlpha{} packages. In Chapter~\ref{basicfunctions}, we present
the main package of \MMAlpha{}. 
Chapter~\ref{domainfunctions} concerns
functions on domains. In Chapter~\ref{staticfunctions}, functions 
related to the static analysis of \alfa{} programs are described.
etc.

This manual is part of the general documentation of \MMAlpha{}.
This documentation can be accessed through the html hierarchy 
that follows the \MMAlpha{} hierarchy. The origin of this 
hierarchy is
\begin{verbatim}
$MMALPHA/doc/index.html
\end{verbatim}
%$
\section*{Some conventions}
Many functions of \MMAlpha{} operate on the Abstract Syntax Tree (AST) of
an \alfa{} program and return the new transformed AST. For example, 
evaluating \texttt{normalize[sys]} returns a normalized version of 
the \texttt{sys}, which may be any \mma{} expression whose evaluation
is the AST of an \Alfa{} program. The result of this transformation may
be stored in another variable, e.g.:
\begin{verbatim}
x = normalize[sys];
\end{verbatim}
and even manipulated by a \mma{} function.

However, \MMAlpha{} provides another, implicit way of manipulating
an \Alfa{} program. When an Alfa{} program is parsed and loaded
using the \texttt{load} function (see \texttt{?load}), 
it becomes the {\em current \alfa{} 
program} and is stored in a variable 
named \$result. By default, functions operate on \$result and 
modify \$result, in addition to returning the transformed program.
For example, 
\begin{verbatim}
normalize[];
\end{verbatim}
puts in \$result the normalized version of \$result, and 
returns this value. (This is the reason of the \texttt{;} ending
this command, otherwise, \mma{} would print out the AST on the notebook.)
Before modifying \$result, this variable is stored in another variable
named \$program, so that the value of the current AST before 
the normalization is kept. The \texttt{undo[]} command allows 
\$result to be restored to its previous value.

\section{Content of the Reference Manual}
The reference manual is structured in the following way.
\begin{enumerate}
\item The mail \LaTeX{} file, \texttt{referenceManual.tex}, is located in 
directory:
\begin{verbatim}
../doc/RefefenceManual
\end{verbatim}
It contains an introduction (that you are currently reading) and calls to 
the various documentation packages, which are produced using commands
of the \texttt{Makedoc} package (see~\ref{doc-makedoc} for details.
\item The reference manual contains 
the following
Chapters.
\begin{itemize}
\item Chapter~\ref{basicfunctions} presents the functions contained in the \texttt{Alpha} 
package. 
\item Chapter~\ref{domainfunctions} presents the functions contained in the \texttt{Domlib},
\texttt{Zpol},
\texttt{Visual}, and \texttt{Visual3D}
packages. 
\item Chapter~\ref{staticfunctions} presents the functions contained in the \texttt{Static},
and \texttt{Semantics}
packages. 
\item Chapter~\ref{subsystemfunctions} presents the functions contained in the \texttt{SubSystems},
package. 
\item Chapter~\ref{manipfunctions} presents the functions contained in the \texttt{Matrix},
and \texttt{Tables}
packages. 
\item Chapter~\ref{elementaryfunctions} presents the functions contained in the \texttt{ChangeOfBasis},
\texttt{Cut},
\texttt{Normalization}, 
\texttt{Reduction}, 
\texttt{Substitution}, 
and \texttt{Substitution}
packages. {\em This chapter has not been proof-read.}
\item Chapter~\ref{scheduling} presents the functions contained in the \texttt{ScheduleTools}, 
\texttt{VertexSchedule}, 
and \texttt{FarkasSchedule}
packages. {\em This chapter has not been proof-read.}
\item Chapter~\ref{elementaryfunctions} presents the functions contained in the \texttt{Pipeline},
\texttt{Control},
and \texttt{PipeControl}
packages. {\em This chapter has not been proof-read.}
\item Chapter~\ref{uniformization} presents the functions contained in the \texttt{Pipeline},
\texttt{Control},
and \texttt{PipeControl}
packages. 
\item Chapter~\ref{backend} presents the functions contained in the \texttt{ToAlpha0v2},
\texttt{Alphard},
\texttt{Vhdl2},
and \texttt{VhdlTestBench}
packages. 
\item Chapter~\ref{utilities} presents the functions contained in the \texttt{MakeDoc},
package. 
\end{itemize}
\item The status of the documentation contained in each package 
is described in a note at the beginning of the package: when the documentation has been
revised, it appears in a note following the subsection of the package. A revision
means that I checked that the content of the usage and note statements of the
package appear correctly on the reference manual, but does not mean that the
documentation itself is correct.
\end{enumerate}
\newpage
\section*{Acknowledgments (Patrice Quinton)}
This version of \MMAlpha{} was developed over the years by the contribution of
many people. With the risk of forgetting some contributors (sorry for this), 
here is a list of these contributors:
\begin{itemize}
\item Christophe Mauras invented the \Alfa{} language in his PhD thesis
en 1989, and the actual language is almost its initial version.
\item The late Herv\'e Le Verge designed with Christophe the first 
version of \Alfa{} (called \Alfa{} du Centaur, as it was designed 
using the Centaur system of Inria.) Herv\'e also designed the first
version of the Polylib which is the engine of \MMAlfa{}.
\item Doran Wilde designed the first version of \MMAlfa{}, i.e., 
the initial \mma{} version of our tool. It started in 1989. 
\item Zbignew Chamski participated to the development of 
this first version. 
\item Tanguy Risset was, and still is, the main architect and 
contributor of \MMAlfa{}. 
\item Florent Dupont de Dinechin implemented a significant part of 
\MMAlfa{} (subsystems, semantic analysis, and part of the back end 
process).
\item Patricia Le Moenner implemented the first version of the 
Vhdl translator.
\item Anne-Claire Guillou contributed to the current Vhdl translater, 
and designed several circuits.
\item Sanjay Rajopadhye was a deep inspirator of \MMAlfa{}.
\item Fabien Quiller\'e implemented mainly the C code generator.
\item Several indian students contributed to the development of
\MMAlfa{} during internships.
\end{itemize}

I was often asked why \MMAlfa{} was developed using \mma{}. The 
choice was made in 1989, mainly in order to take advantage of the 
available formal calculations embedded in this tool. There are
disadvantages to this choice, one being that people have
to buy this software in order to use \MMAlfa{}. I believe that this
is not the main point, however, since \mma{} is available in many
universities, and companies interested in our software could 
buy it easily. Probably more important is that \mma{} does not fit in 
the current {\em culture} of developers.

But from the point of view of a prototype developper, \mma{} is 
a very nice tool. Our software is available on {\em all platforms}
thanks to the portability of \mma{}\footnote{By the way, the 
part of \MMAlfa{} that creates most portability problems is 
the Domlib library that is written in C.}. I have never experienced 
any problem of running out of space, and time constraints have 
never been a significant issue. Moreover, every 2 years or so, 
\MMAlfa{} accelerates significantly when I change my laptop\footnote{The only
occasion when this was not true is when I recently changed my Apple MacBook 
into a MacBook Pro, where I was disappointed to see the performance
degrade. But it may be only because I did not recompile Domlib...}.

\chapter{Basic Functions}
\label{basicfunctions}

\section{Introduction}
Basic functions are mainly in the Alpha package.
\input{Alpha}

\chapter{Functions on domains}
\label{domainfunctions}
\input{Domlib}%Proof-read
\input{Zpol}%Proof-read
\input{Visual}%Proof-read
\input{Visual3D}%Proof-read

\chapter{Static Analysis}
\label{staticfunctions}
\input{Static}
\input{Semantics}

\chapter{Subsystems}
This package contains functions to handle Alpha subsystems.
\label{subsystemfunctions}
\input{SubSystems}

\chapter{Manipulating Alpha Expressions}
\label{manipfunctions}
The Matrix and Tables packages contain some functions to manipulate
Alpha expressions and programs. Other such functions are scattered
in other packages, as they have been written sometimes on the fly
by the authors of these packages. A developper has to read this
chapter carefully in order not to reinvent the wheel.
\input{Matrix}
\input{Tables}

\chapter{Elementary Program Transformations}
\label{elementaryfunctions}
\input{ChangeOfBasis} %Proof-read
\input{Cut} %Proof-read
\input{Normalization} %Proof-read
\input{Reduction} %Proof-read
\input{Substitution}

\chapter{Scheduling}
\label{scheduling}
\input{ScheduleTools}
\input{Schedule}
\input{VertexSchedule}
\input{FarkasSchedule}

\chapter{Uniformization}
\label{uniformization}
\input{Pipeline}
\input{Control}
\input{PipeControl}

\chapter{Back end process}
\label{backend}
\input{ToAlpha0v2}
\input{Alphard}
\input{Vhdl2}
\input{VhdlTestBench}

\chapter{Utilities}
\label{utilities}
\input{MakeDoc}%Proof-read

\chapter{Add-ons}
\input{BinExpansion}
\input{BitOperators}
\input{Decomposition}
\input{DesignManagement}
\input{INorm}
\input{INormalize}
\input{Lexicographic}
\input{LinearAlg}
\input{Matrix}
\input{Meta}
\input{ModularSchedule}
\input{Omega}
\input{Options}
\input{Properties}
\input{Reduction}
\input{Schematics}
\input{Semantics}
\input{Static}
\input{SubSystems}
\input{SystemCTestBench}
\input{SystemCTools}
\input{Tables}
\input{Transformation}
\input{Uniformization}
\input{UniformizationTools}
\input{Visual}
\input{Visual3D}
\input{Zpol}

\printindex

\end{document}