### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Schedule/Scheduler_dev_man.tex

\documentstyle[fullpage]{article}

\newcommand{\aalpha}{{\sc Alpha}}
\newcommand{\ammalpha}{{\sc MmAlpha}}
\newcommand{\err}{\top}
\newcommand{\indef}{\bot}
\newtheorem{theorem}{theorem}

\title{reference manual for the Alpha {\bf Scheduler}}
\author{Tanguy Risset}
\date{\today}

\begin{document}
%\bibliographystyle{plain}

\maketitle

This documentation provides help for the user of the {\aalpha} 
scheduler developer. It is available in the file 
{\tt \$MMALPHA/doc/developper/Scheduler\_dev\_man.ps}

\section{Introduction}

 The basic goal of the scheduler 
is to find a valid execution order that is {\em good} with respect to 
a particular criterion. The theoretical basis for the scheduling process 
is inherited from the systolic array synthesis researches and from the 
automatic parallization researches. Our  implementation makes use of the 
procedure defined in~\cite{Feautrier92aa}. Rouglhy speaking, 
the time is considered as a discrete  
single clock. the idea of the scheduling process is to 
gather all the constraints that must verify the schedule in a 
 linear programming problem (LP) and to solve it with a particular 
software.

This document is not intended to be a comprehensive explanation of the
implementation, it is an introduction to the overall architecture of
the package {\tt NewSchedule.m} which is quite big (about 3000 lignes).
The developper should carefully read this before getting into the
code.  The source code of {\tt NewSchedule.m} is commented in such a way
that the present document document should be enough to understand
everything.

\section{theoretical basis}
\label{theoric}
The schedule is based on the method proposed in~\cite{Feautrier92aa}
and uses the farkas lemma. We give her the
form of this lemma that we use for ensuring constraints
\begin{theorem}[Affine form of Farkas lemma]
\label{affineFeautrier}
Let ${\cal D}$ be a polyhedron defined by the inequalities $Ax + b \geq 0 $,
Then the affine form $\phi$ is non-negative on ${\cal D}$ if and only
if there exists $\lambda_0,\Lambda \geq 0$ such that $\phi=\Lambda
\left(\begin{array}{cc} A & b \end{array}\right) +\lambda_0$
\end{theorem}
{\bf Proof:} See~\cite{Schrijver86} .

It is easy to see how we will make use of this lemma. To ensure a linear 
constraint (which can always be written as the fact that a linear function is positive on a domain), we will try to find $\Lambda,\lambda_0$ such that 
the above equality is verified (we transformed a $\forall x \in ...$ in 
a $\exists y\geq 0 s.t....$).  The method proposed 
in~\cite{Feautrier92aa} distinguish
three type of constraints that a  scheduling function $T$ must ensure. 
\begin{itemize}
\item {\bf Positivity constraints}: for each variable $A$ of the program, for any 
point $(x,N)$ in the domain of $A$, we must have $T_A(x,N) \geq 0$.
\item {\bf Dependency constraints}: for any 
dependence $A[x,N] = F(\ldots,B[I(x,N)], \ldots)$, for each point $(x,N)$ 
where this dependence occurs, we must have: 
$T_A(x,N) - T_B(I(x,N)) - \delta \geq 0 $ 
(here $\delta$ is a delai associated to 
the computation of $A$.
\item {\bf Objective function constraints}: These constraints depend on the type of 
objective function choosen. If we chose to minimiz
e the overall execution time, 
we introduced a function $f(N)$ depending only upon the parameters of
 the program, which will be greater than any time scheduled (i.e. for each variable $A$,  for any 
point $(x,N)$ in the domains of $A$, we must have 
$f(N) - T_A(x,N) \geq 0$).
\end{itemize}
As index functions are affine functions
 in \aalpha, all these constraints are of 
the required form: ensure that an affine function is positive on a particular 
domain, thus we can use  the Farkas lemma. 
For each  constraints, we will use the formulae of 
theorem~\ref{affineFeautrier} and 
 obtain a system of equalities that will be solve with mppip.
\begin{itemize}
\item Positivity constraints: if a variable $A$ is defined on the domain
$D_A=\{x ~|~ Cx + b \geq 0 \}$ ($x$ has $n$ components and $C$ has 
$k$ rows), the scheduling function 
$T_A(x) = Tx+ t$ is positive on $D_A$ if and only if the system: \\
\centerline{
$\left\{\begin{array}{l} (T\ t) - \Lambda.(C\ b) - \lambda_0 = 0   \\
\lambda_i \geq 0, 0 \leq i \leq k\end{array}\right.$
}
has a solution (where $\Lambda$ is a $k$-vectors whose $i^{th}$ component 
is $\lambda_i$). Hence we get $n+1$ equation: 
\begin{itemize}
\item for each dimension $i$, if $C_{\cdot i}$ is the $i^{th}$ 
column of C:  $\tau_i   - \Lambda.C_{\cdot i} = 0$
\item $t -  \Lambda.b - \lambda_0 = 0$
\end{itemize}

\item dependency constraints: $A[x]$ depends upon $B[I(x)]$ where 
$I(x) = Dx+d$ on the domain $D_{AB} = \{x ~|~ Cx + b \geq 0 \}$
($x$ has $n$ components and $C$ has $k$ rows) the scheduling functions 
$T_A(x) = T_Ax+ t_A$ and  $T_B(x) = T_Bx+ t_B$ respect the dependency 
if an only if the system:\\
\centerline{
$\left\{\begin{array}{l} (T_A\ t_A) - (T_B \ t_B).(D \ d) -  \delta_{BA} - M.(C\ b) - \mu_0 = 0   \\
\mu_i \geq 0, 0 \leq i \leq k\end{array}\right.$
}
has a solution (where $M$ is a $k$-vectors whose $i^{th}$ component 
is $\mu_i$, $\delta_{BA}$ is the delai necessary between the production of $B$ and the production of $A$). Hence we get $n+1$ equation: 
\begin{itemize}
\item  for each dimension $i$,  $\tau_{A,i}   - (T_B \ t_B).D_{\cdot i} - 
M.C_{\cdot i} = 0$
\item $t_A -  (T_B \ t_B).d -\delta_{BA}- M.b - \mu_0 = 0$
\end{itemize}

\item  Objective function constraints: 
if a variable $A$ is defined on the domain
$D_A$ with $D_A=\{(x,N) ~|~ A\left(\begin{array}{c} x\\ N \end{array}\right) + b \geq 0 \}$ ($x$ has $n$ components, $N$ is the parameter vectors has $m$ component and $A$ has 
$k$ rows), the scheduling function 
$T_A(x) = T_A\left(\begin{array}{c} x\\ N \end{array}\right) + t_A$ 
lower than the objective function $\Phi.N + \phi_0$ on  $D_A$ 
if and only if the system: \\
\centerline{$\left\{\begin{array}{l} (0 \ \Phi \ \phi_0) - (T_A\ t_A) - R.(A\ b) - \rho_0 = 0   \\
\rho_i \geq 0, 0 \leq i \leq k\end{array}\right.$
}
has a solution (where $R$ is a $k$-vectors whose $i^{th}$ component 
is $\rho_i$).  Hence we get $n+m+1$ equation: 
\begin{itemize}
\item for each $i, \ 1 \leq i \leq n$:  $-\tau_{A,i}   - R.A_{\cdot i} = 0$
\item for each $i, \  n+1 \leq i \leq n+m$: $\phi_{i-n} -  \tau_{A,i}   - R.A_{\cdot i} = 0$
\item $\phi_0 - t_A - R.b - \rho_0 = 0$
\end{itemize}
\end{itemize}

\section{Algorithm of the schedule function}

The basic structure of the fonction schedule is the following:
\begin{enumerate}
\item Perform various analysis (get the list of variable, domains for each 
 variables,  dependencies, duration for each instruction \ldots)
\item Build the list of constraints of the linear program (three types 
of constraints: constraints ensuring positivity for timing function,
constraints  ensuring dependancies, constraints for the objective function).
\item Write the LP in a unix file in {\tt Pip}. Solve the LP with 
the software  {\tt mppip}, and get the result.
\item Print the result and get the final output form.
\end{enumerate}

\section{Step 1: various analysis}
\begin{itemize}
\item {\tt addAllParameterDomain}. The syntax of parameter in Alpha allow 
some context constraints that are implicit in all domains of the \aalpha 
program.  
If we want to apply Farkas lemma, we need all the constraints of the domain, 
hence, we add {\em by hand} these constraints to all domains of the program.

\item {\tt domains of variables}. We have choosen to distinguish three
types of variables: input variables, local variables and Output
variables. For each variable, the domain is convexise, i.e. we take
the convex hull of the real domain of the variable (this allow to
simplify the resulting LP, but imply a less powerfull scheduler).

\item {\tt dependencies}. The dependence analysis is first done 
by the {\tt dep[] } function. Then,  the list of dependancies are 
classified according to the type of dependancy (6 type of dependency, in this 
order: Local variable 
to Output variable, Input to Output, Output to Local,  Input to Local,
Local to Local, Output to Output).

\item {\tt makeNumInstr}. This function numbers all variable. The number 
of a variable is its order of declaration in the Alpha program. 
This function build a list of couple: {\tt  \{name\_String,number\_Integer\}}
 that will be used 
in the rest of the program to make the correspondance between
 the name of a variable and its number.

\item {\tt makeRefline}. At this point of the program, we are able to
know exactly how many variables will have the linear program generated
(how many $\lambda$'s, $\rho$'s, etc...). As we will see below, we
will first build a constraint matrix in which each row has the same
structure.  This structure is indicated by the {\tt refline}
constructed here.  {\tt refline} contains 8 sublist.  The first one
indicates how many coefficient there are in the objective function
$\Phi$. The second one indicates, for each variable $A$, the dimension
of the domain of $A$ (hence the number of $\tau_A$'s). The third one
indicates how many variables will be dedicated to the constant part of
the scheduling functions (the $\alpha$'s). The fourth indicates, for
each dependancy, how many farkas coefficient ($mu$) will be introduced
by the equation needed to ensure this dependancy. The fifth indicates,
for each variable, how many farkas coefficient ($\lambda$)
 will be necessary to 
ensure the positivity condition for this variable, The sixth indicates, 
for each variable, how many farkas coefficient ($\rho$)
 will be necessary for the objective function  constraints. the $7^{th}$ and 
$8^{th}$ elements are set to {\tt \{1\}} and will be explained later.

As we have classified the variables in three categories and
the dependancies in 6 categories, elements $2,3,5,6$ of refline are
divised in three sublist and element $4$ of refline is divised in 6
sublist.  The precise structure of {\tt refline } is clearly commented
in the {\tt NewSchedule.m } package.  For the moment, we just  need to 
know that this variable will be used in the rest of the program for 
getting information on the program (for instance: how many input variable, 
how many $\lambda$'s related to variable $A$ etc.).

\end{itemize}

\section{Step 2: building the list of constraint}
the list of constraints is built by the function {\tt makeTableEq}.
As we have explained in section~\ref{theoric}, there are three types
of constraints. The positivity constraints are set for the Input,
Output and Local variables, but we impose the objective function
constraint only on the output variables. the dependency constraints
are built for the six type of dependencies. The constraints are the
elements of the {\tt structTableEq}. Each constraint is a list which
have the following structure: \\
\begin{verbatim}
{ PhiCoeffs_IntList,
{ TauLocal_IntListList,TauOutput_IntListList,TauInput_IntListList},
{ AlphaLocal_IntList,AlphaOutput_IntList,AlphaInput_IntList},
{ LambdaLocToOut_IntListList,  LambdaInToOut_IntListList,
  LambdaOutToLoc_IntListList,  LambdaInToLoc_IntListList,
  LambdaLocToLoc_IntListList,  LambdaOutToOut_IntListList},
{ MuLoc_IntListList,  MuOut_IntListList,  MuIn_IntListList },
{ RhoLoc_IntListList,  RhoOut_IntListList,  RhoIn_IntListList },
{ CstCoeff_Int },
{ EqOrIneq_Int} }
\end{verbatim}
In the above structure, {\tt \_IntList} indicates that the element considered
is a list of integer (and {\tt \_IntListList}, a list of list of integer). 
In a constraints {\tt c1} of this type, 
{\tt PhiCoeffs} is the list of factors of the 
coefficients of the objective function $\Phi$ in {\tt c1}.
 One element of {\tt TauLocal} correspond to a local variable (say $A$) 
and gives the factor (in c1)  of the components 
of the scheduling vectors of $A$. 
One element of {\tt LambdaLocToOut} corresponds to a particular 
dependency from a local variable to
 an output variable and gives the coefficients (in c1)
 of the  farkas coefficient $\lambda$ introduced for this dependency,
 and so on ....

For instance, if $A$ is the first local 
variable of the Alpha program which has only one parameter 
 (the domain of $A$ has dimension 3),
 and we want to impose the constraint: 
$\phi_1 - 2\tau_{A,1} = 0 $, this constraint we be stored in one row as:
\begin{verbatim}
{ {1,0},{{{-2,0,0},{0.... (* structured list of  0 in the remaining *) ,{0}}
\end{verbatim}
and the real constraint can be obtained by 
$expr1 = (1\ 0).\left(\begin{array}{c} \phi_1 \\ \phi_0 \end{array}\right)+
 (-2\ 0 \ 0).\left(\begin{array}{c} \tau_{A,1} \\ 
\tau_{A,2} \\ \tau_{A,3}  \end{array}\right)+ \ldots  = 0 $.

In {\tt NewSchedule.m}  such structured  constraints are always assigned to 
variables which are called {\tt line0} or {\tt line1} (like the lines of the 
constraint matrix). Usually, {\tt line0} consists of a structured constraints
corresponding to 0=0 (build from the informations contained in {\tt refline}),
 then it is assigned to {\tt line1}, and
 the  components of {\tt line1} are iteratively modified to obtain a 
valid constraint as indicated in section~\ref{theoric}.
 The $7^th$  element of  line1 is a list of one single integer
which is the constant that appear in the constraint. The $8^th$  
 element of  {\tt line1} is a list of one single integer which indicates if the constraints is and equality (value 0) or an inequality (value 1: $expr1 \geq 0$)

\section{Step 3: Solving the LP}

{\tt fichierPip}  writes the constraints of 
{\tt structTableEq} matrix in the mppip format in a file, 
then calls {\tt mpppip} to solve the LP and read the solution.

{\tt fichierPip} first calls {\tt matPip} which flatten the output of
each line of {\tt structTablEq}. At this point, the order of the coefficients 
of the $\tau$'s is inverted in order to minimize first the coefficients 
of the parameters. {\tt matPip} outputs the matrix {\tt matrice} 
 Then, {\tt writefile} writes the matrix in a file (this part is
quite long for big files). During this phase we add a single parameter
to the LP which is specific to {\tt Pip } behaviour. As we want to allow some 
variables to be negative (the $\tau$'s for instance), we have to add a 
{\em big parameter} $G$ and replace each occurrence of a $\tau$ by $\tau-G$. 
This manipulation is explained in~\cite{Feautrier88a}, This should be transparent 
and the coefficient of the big parameter in the solution is not important.

After the matrix is written, we perform some unix manipulations on the
file to obtain the good mppip input format. Then, we run 
mppip~\cite{Feautrier92aa}  and
the result is read by {\tt readPipResult},

\section{Step 4: final output form}
{\tt reorderSchedResult} place back the coefficients of the parameters
at the end of the scheduling vectors.
 {\tt getFinalOutputForm} pretty prints the result to the screen and 
set the output format to the following form: \\
{\tt <schedResult>=Alpha`ScheduleResult[scheduleType\_Integer,<sched3List>]\\
<sched3>=$\begin{array}{l} \mbox{\{ nameVar\_String,}\\
          \mbox{ indices\_List,}\\
           \mbox{ Alpha`sched[tauVector\_List,constCoef\_Integer] \}}
\end{array}$
}

 \bibliographystyle{alpha99}
\bibliography{/udd/risset/TEX/ref/newbib}
\end{document}