### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/appendixtechset.tex

\chapter{Installing \mmalfa{}}
\label{appendixtechset}
\label{install}

\newcommand{\mmalpha}{{MMalpha}}
\input{install1.tex}

\section{Getting started}
You may use either {\mma} under {\tt emacs}, or directly on the Unix 
system. 
\begin{itemize}
\item under {\tt emacs}, type {\tt  esc-X Mathematica}. 
\item type math (use the {\mma} textual interface).
\item type mathematica (use the {\mma} textual interface).
\end{itemize}
%Do not use a Notebook, as the output of the pretty printer would not
%be available.
When the {\mma} session is started, you can type {\tt goDemo["INIT"]}
for the initiation demonstration.

When the {\mma} session is started, you can type {\tt on[]} or
\texttt{start[]} for the initiation demonstration.

\section{{\mmalpha} file hierarchy}
\includegraphics[width=12cm]{figures/fileHierarchy}