### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/chapalphascal.tex

\chapter{Modelling synchronous architectures}
\label{chapalphascal}

\begin{tobedone}
\begin{enumerate}
\item Redraw figures (after reinstalling X11 and xfig)
\end{enumerate}
\end{tobedone}

In this chapter, we present the {\alfa} language through the 
description of very simple -- and seemingly artificial -- examples
of synchronous circuits\index{synchronous circuits}. Our goal is to show the basic constructors
of the language: their syntax, their meaning, and how they can be
used in practice. 
% 
%\newpage
\section{Pointwise operators\index{pointwise operators}}
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
%\colorbox{white}{
\centerline{\includegraphics{figures/exemple_1}}
\caption{An adder\label{adder1}}
\end{figure}
Consider the architecture\index{architecture} shown in figure~\ref{adder1}: there is a 
simple adder\index{adder}, 
which takes to input flows, respectively {\tt x} and 
{\tt y}, and returns {\tt S}.

This architecture is described by the {\alfa} program of figure \ref{adder-alpha}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\verbatiminput{alpha_progs/adder.alpha}
\caption{An adder in {\alfa}}\label{adder-alpha}
\end{figure}
A program is called a {\em system}\index{system}. 
Line 1 is the head of the system, 
with name {\tt exemple1}. This system has two input arguments
named
{\tt x} and {\tt y}. These arguments, called {\em variables}\index{input
variables}, are
defined on the set {\tt \{t | 1 <= t\}}. Line 3 defines a local variable\index{local variable}
{\tt S} (for the moment, this variable is useless, but we shall
soon find a use for it). 
Between the {\tt let} and {\tt tel} keywords of lines 4 and
7, we have two equations\index{equation}: 
line 5 says that {\tt S} is the pointwise sum of 
{\tt x} and {\tt y}. In other words, $\forall t, S_t = x_t + y_t$. 
Line 6 says that {\tt z} and {\tt S} are identical.
\subsection*{Processing this example in {\mmalfa}}
Here are the Mathematica expressions that allow one to try
the previous example under MMAlpha.
%A more complete introduction to {\mmalfa} is given in 
%appendix~\ref{mma-introduction}.
\begin{verbatim}
load["adder.alpha"]    (* loads the alpha file *)
ashow[];               (* shows its content *)
\end{verbatim}

%
%\newpage
\section{Delays as dependence fonctions}
We now modify this example, by adding a delay\index{delay} on the output, as
shown in figure \ref{adderDelayed}. 
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exemple_2}}
\caption{Adder delayed}\label{adderDelayed}
\end{figure}
The corresponding {\alfa} program is shown in figure \ref{adderDelayed-alpha}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\verbatiminput{alpha_progs/adderDelayed.alpha}
\caption{Adder plus delay}\label{adderDelayed-alpha}
\end{figure}
In this example, the definition domains\index{domain} of {\tt z} and {\tt S} have
changed: indeed, the output of the adder\index{adder (delayed)} is defined only one
cycle after the input. Thus, {\tt z} and {\tt S} are declared on 
domain {\tt \{t|2<=t\}}. Instruction 4 now defines {\tt S} as being the
sum of variables {\tt x+y}, delayed by 1.

%
%\newpage
\section{Retiming\index{retiming}}
{\alfa} provides tools for retiming synchronous architectures. 
In the example of figure~\ref{adderDelayed}, one can shift the 
delay from the output of the adder to its inputs. 
This corresponds to replacing line 4 in program~\ref{adderDelayed-alpha}
by \verb\S[t] = x[t-1] + y[t-1];\. The corresponding architecture is shown 
in figure~\ref{adderRetimed}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exemple_3}}
\caption{Retiming the adder}\label{adderRetimed}
\end{figure}
\subsection*{Retiming in {\mmalfa}}
Do 
\begin{verbatim}
load["adderDelayed.alpha"]; (* load the alpha program *)
ashow[];                      (* show it *)
normalize[];                  (* simplify it *)
ashow[];                      (* show the result *)
asave["file.alpha"];          (* save program *)
\end{verbatim}

%
%\newpage
\section{Case and restrictions}
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exemple_5}}
\caption{Adding a multiplexer}\label{adderMultiplexer}
\end{figure}

Two other constructs of the language are the case\index{case} and the
restriction.  Consider the example of figure~\ref{adderMultiplexer},
which shows a new delayed and retimed version of the adder, where the
output is taken through a multiplexer\index{multiplexer}. In other
words, depending on the value of index {\tt t}, {\tt S} is either 0 or
the sum {\tt x[t-1] + y[t-1]}. This is reflected in the definition of
{\tt S} in figure~\ref{addermultiplexer-alpha} (lines 2--5.) This
definition uses a case expression. Each branch\index{branch (of a case
expression)} of the case is an {\alfa} expression. The first one is
the constant\index{constant} $0$, denoted as {\tt 0[]} in {\alfa}. It is
considered only when {\tt t} is equal to $1$, which is reflected in the
condition\index{condition} {\tt \{| t = 1\}}. The second branch
corresponds to the expression {\tt x[t] + y[t]}, taken when the
condition {\tt \{|t>1\}} is true. Actually, we call an expression such
as {\tt \{|t>1\}} a {\em restriction}.\index{restriction}

\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\verbatiminput{alpha_progs/adderMultiplexer.alpha}
\caption{Adder and multiplexer}\label{addermultiplexer-alpha}
\end{figure}
The definition of {\tt z} also uses a restriction. The definition of 
line 6 says that {\tt z} is defined as {\tt S}, but only for time 
instants {\tt t} such that {\tt \{|t>=5\}}. 
\subsection*{Exercices in {\mma}}
Load and parse program adderMultiplexer.alpha.
\begin{verbatim}
load[adderMultiplexer.alpha];
ashow[adderMultiplexer];
\end{verbatim}

%
%\newpage
\section{Array and full form of {\alfa}}
\label{arrayform}

In all the above examples, we have used the so called {\em array
form}\index{array form} of {\alfa}. The full form\index{full form} of
the language is actually a little bit different, and a little bit more
difficult to read initially. It is however needed in order to
understand some of the properties of the language.

Let us consider again the definition of {\tt S} in
program~\ref{addermultiplexer-alpha}, line 2 -- 5:
\begin{verbatim}
 S[t] = case
     {|t=1}: 0[];
     {|t>1}: x[t-1] + y[t-1];
     esac;
\end{verbatim}
The full form of this expression is in fact: 
\begin{verbatim}
 S = case
      {t|t=1}: 0.(t->);
      {t|t>1}: x.(t->t-1) + y.(t->t-1);
  esac;
\end{verbatim}
To explain this, we need to go more deeply in the 
structure of the language. In {\alfa}, any expression is a function from
a set of points -- called the {\em domain} of the expression --, 
to some value set depending on the type of the expression. 
Expressions are either 
\begin{itemize}
\item variables -- e.g. {\tt S},\index{variables} 
\item scalar constants\index{constants}
\item pointwise operations on expressions -- e.g. {\tt x + y}\index{pointwise 
operators}
\item application of dependence functions to expressions, -- e.g. 
{\tt exp.f} where {\tt f} is a dependence function explained later,\index{dependence function}
\item restriction of expressions, -- e.g. {\tt dom: exp}, where 
{\tt dom} is a domain description\index{restriction},
\item case expressions -- e.g. {\tt case e1 ; e2; esac}, where {\tt e1}
and {\tt e2} are expressions.\index{case expression}
\end{itemize}

Let us examine expressions in more details. 
\subsection{Domains}
\index{domain}

Domains of {\alfa} are restricted to {\em polyhedra}, (more precisely,
unions of finitely may polyhedra).  A polyhedron is a set of integral
points whose coordinate satisfy linear inequalities.  For example:
{\tt \{t | 1 <= t <= 10\}} denotes such a set. So also does {\tt \{i,j
| i>=10; 0 <= j <= i\}} for example. {\texttt \{i | 0 < i < 10\} ||
\{i | 15 <= i\}} is a union of polyhedra.  \index{polyhedron}

\subsection{Variables}
\index{variable}
Variables are functions from a domain to a set of values. For example, 
in program~\ref{addermultiplexer-alpha}, {\tt S} is a function from 
the domain {\tt \{t|1<=t\}} to the set of integers. Thus, {\tt S}
represents a collection of indexed values $S_t$, where {\it t} belongs
to the domain of {\tt S}.

{\alfa} is a strongly-typed language, where variables are declared,
together with their domain. This allows many compile time checks to be
performed\index{type-checking} (see chapter~\ref{chapanalyze}).

Input variables are defined in the head of the system
definition. Output variables are defined after the key-word {\tt
result}. Local variables are defined after the key-word {\texttt var}
and the keyword {\tt let}.

\subsection{Constants}
In {\alfa}, constants are considered as functions from the
singleton set $Z^0$ to a type. For example, $1$ is the function 
from $Z^0$ to the integer value $1$. 
\subsection{Pointwise operations}
\index{pointwise operations}

Pointwise operations generalize classical operations like +, -, * on
integers, or {\tt and}, {\tt not}, {\tt or} on booleans. So, {\tt x +
y} represents the pointwise operation on values $x_t$ and $y_t$ for
all $t$. There is however, an important point to notice: if {\tt op}
is a pointwise operator and {\tt e1} and {\tt e2} are expressions,
then the domain of the expression {\tt e1 op e2} is the {intersection}
of the domains of {\tt e1} and {\tt e2}.\index{domain of a pointwise
expression}

\subsection{Dependence functions}
\index{dependence function}

Dependence functions are affine functions on indices.
They are denoted as
\begin{center}
\texttt{(idx, idx, ... -> iexp, iexp, ... )}
\end{center}
where \texttt{idx} is an index name and \texttt{iexp} is an affine
function of indices.  For example: {\tt (i,j -> i+2, j+3i-4)\}}
denotes such a function.  So does also the function {\tt f=(t ->
t-1)}. Therefore, {\tt x.(t -> t-1)} is an {\alfa} expression, whose
meaning is as follows.  Remember that {\tt x} is a function from the
domain {\tt \{t | 1<=t\}} to the set of integers. The composition of
functions {\tt x o f} is therefore defined, and this is just what {\tt
x.(t -> t-1)} represents. It is easy to see that this function is
defined on the domain {\tt \{t | t>=2\}}.  More generally, the domain
of any expression {\tt e.f} is ${\tt f}^{-1}(\mbox{dom}({\tt e}))$,
that is to say, the pre-image of the domain of {\tt e} by function
{\tt f}.

Note that a function {\tt (->t)} denotes a mapping from the set $Z^0$
to $Z$. Similarly, {\tt (t ->)} denotes a mapping from the set $Z$ to
$Z^0$. Such functions are used to extend scalar constants to higher
dimensional domains.\index{Extension of constants}

\subsection{Restriction expressions}

\index{restriction} As said before, restrictions allows one to
restrict the definition of an expression to a subdomain. To be
consistent with the other language constructs, restrictions are
defined as polyhedra, and therefore are denoted just as domains. For
example, {\tt \{t | t>=5\}: S} is a restriction of the expression
\texttt{S} to domain \texttt{\{t | t >= 5\}}. The domain of a
restricted expression \texttt{dom: e} is the intersection of the
domain of the expression and the domain defined by the restriction,
i.e., $\texttt{dom} \cap \texttt{e}$.\index{domain of a restriction}

\subsection{Case expressions}

\index{case expression} As already seen, case expressions allow one to
define a new function case by case. A case expression is defined on
the union of the domains of its branch. Let {\tt case exp1; exp2 esac}
be a case expression, with branches {\tt exp1} and {\tt exp2}. Let
{\tt d1} and {\tt d2} be respectively the domains of {\tt exp1} and
{\tt exp2}. Assume that {\tt d1} and {\tt d2} have the same
dimensionality and are disjoint (these properties are checked during
static analysis). Assume also that {\tt exp1} and {\tt exp2} have the
same type, say {\tt T}.  Then {\tt case exp1; exp2 esac} is a function
from ${\tt d1} \cup {\tt d2}$ to {\tt T}, whose value is that of {\tt
exp1} on {\tt d1}, and that of {\tt exp2} on domain {\tt d2}. For 
a detailed desciption of the domain computations and its use for the
static analysis, see section~\ref{static-analysis}.

\subsection*{Back to our example}
We are now ready to consider again the example corresponding to figure~\ref{addermultiplexer-alpha}. The definition of {\tt S} is:
\begin{verbatim}
 S = case
      {t|t=1}: 0.(t->);
      {t|t>1}: x.(t->t-1) + y.(t->t-1);
  esac;
\end{verbatim}
Let us analyze each branch. {\tt 0} is a function from $Z^0$ to 
the integer value $0$. Composing this function with the function 
\texttt{(t ->} extends the constant {\tt 0} to the set $Z$. In other 
words, {\tt 0.(t->)} denotes the function from $Z$ to integers,
whose value is {\tt 0} everywhere. 
{\tt \{t | t=1\}: 0.(t->)} denotes the restriction of this function 
to the single point {\tt t=1} of $Z$.

We can analyze the second branch in a similar fashion. We would conclude
that it represents a function from the subset {\tt \{t | t>1\}} of $Z$, 
whose value at point {\tt t} is $x_{t-1} + y_{t-1}$.

Finally, the total case expression defines {\tt S} as a function from
{\tt \{t | t >=1\}} to the integer set, whose value is {\tt 0} 
when {\tt t = 1}, $x_{t-1}+ y_{t-1}$ otherwise.

\subsection{Array form}
\index{array form}

The array form is just a convenient way of presenting {\alfa} programs, 
in a more ``readable'' form. The rule is simple to explain on the 
definition of {\tt S}:
\begin{itemize}
\item Dependence functions of the form {\tt .(indices -> exprs)} are rewritten
as {\tt [exprs]}, 
\item Restrictions of the form {\tt \{indices | inequalities\}} are rewritten
{\tt \{| inequalities\}},
\item Left-hand side variables are appended with the indices from the
right hand side.
\end{itemize}
Thus, the definition of {\tt S} in array form is
\begin{verbatim}
 S[t] = case
      {|t=1}: 0[];
      {|t>1}: x[t-1] + y[t-1];
  esac;
\end{verbatim}
This notation can be used safely for any normalized expression, as 
we shall see in section~\ref{substitution}.
\subsection*{Exercices in {\mmalfa}}
Load a program, and see its full form using the {\tt show[]}
function, or its array form using {\tt ashow[]}.
%\newpage
\section{Substitution and normalization}
\label{substitution}
One of the properties of {\alfa} is called the referential transparency\,:
\index{referential transparency}
it means that the meaning of an expression is independent of the context
where it appears. 
For this reason, any symbol can be safely replaced by its definition. 
For example, by replacing the {\tt S} symbol in the definition of 
{\tt z} in program~\ref{addermultiplexer-alpha}, we obtain the 
program shown in figure~\ref{addermultiplexer-alpha}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\verbatiminput{alpha_progs/adderSubstituted.alpha}
\caption{Program of figure~\ref{addermultiplexer-alpha} after substitution of {\tt S} in 
definition of {\tt z}}\label{addersubstituted-alpha}
\end{figure}

Let us consider the new definition of {\tt z}:
\begin{verbatim}
z = {t | 5<=t} : 
         case
           {t | t=1} : 0.(t->);
           {t | 2<=t} : x.(t->t-1) + y.(t->t-1);
         esac.(t->t);
\end{verbatim}

\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\verbatiminput{alpha_progs/adderNormalized.alpha}
\caption{Normalization of program~\ref{addersubstituted-alpha} (in
array form)}
\label{addernormalized-alpha}
\end{figure}

It is rather intuitive that the first restriction {\tt \{t | 5<=t\}}
can be distributed in the branches of the case expression. Combining
(i.e., intersecting) this restriction with the {\tt \{t | t=1\}} leads
obviously to an empty set, and consequently, to an empty
branch. Similarly, combining {\tt \{t | 5<=t\}} with {\tt \{t | 2<=t\}}
leads to {\tt \{t | 5<=t\}}. In summary, we can imagine that the full
definition of {\tt z} can be simplified to something like {\tt z =
x.(t->t-1) + y.(t->t-1)}.

\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exemple_7}}
\caption{Architecture after substitution and normalization}
\label{addernormalized}
\end{figure}
These intuitive properties are in fact true, and such a simplification
is done in a systematic way by means of the normalization
transformation: one can prove that any expression can be simplified
into a normalized one, with at most one level of outermost case, one
level of restrictions, and one level of
dependences.\index{normalization}\index{simplification of expressions}

After normalization, the program becomes that of
figure~\ref{addernormalized-alpha}.

We can interpret this program as representing the synchronous
architecture of figure~\ref{addernormalized}.
We can see -- and this is of course straightforward -- that the 
multiplexer is indeed useless, since we are interested in the 
value of {\tt z} only when {\tt t>1}. This normalization 
property is one of the most important one of {\alfa}.

\subsection*{Exercices in {\mma}}
\subsubsection*{Exercice 1}
Write an {\alfa} program to model the 
following
architecture:
%of figure~\ref{exo1}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exo1}}
\caption{Exercice 1}
\label{exo1}
\end{figure}
\subsubsection*{Exercice 2}
Write an {\alfa} program to model 
the circuit of figure~\ref{exo2}.
\begin{figure}[htbp]
%\htmlimage{scale=\myscale}
\centerline{\includegraphics{figures/exo2}}
\caption{Exercice 2}\label{exo2}
\end{figure}