### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/demos.tex

\chapter{The demos of {\mmalfa}}
\label{demos}
This chapter presents a few demos of {\mmalfa}.\index{demos,demonstrations}
There two types of demos, the ones that are to be seen in a bloc note interface and the ones that should be executed in kernel mode. There are also various documentations available with the archive.

\section{Bloc notes demos}

\begin{description}
 \item [Introduction.] Introduction to {\Alpha}
\item [Domlib.] Elementary computations on domains with Domlib.

 \item [FIR: Design of a FIR filter.]  The FIR demo shows the typical
steps for the synthesis of a linear architecture for a Finite Impulse
Response filter.  The demo comprises the following steps: loading the
initial program, pipelining, scheduling, assign parameter values,
control signal generation, {\alphaz} generation, {\alphard} generation.

 \item [MATVECT: architecture for matrix vector
multiplication.] 
The MATVECT demo shows the typical steps for the
synthesis of a linear architecture for a Matrix Vector Multiplication.
The demo comprises the following steps: loading the initial program,
pipelining, scheduling, assign parameter values, control signal
generation,  {\alphaz} generation, {\alphard} generation.

\end{description}

\section{kernel demos}
These demos are avaiblable at the following address:
\htmladdnormallink{file:/udd/alpha/alpha\_beta/Mathematica/demos/demos.html}
{file:/udd/alpha/alpha\_beta/Mathematica/demos/demos.html}

\begin{description}
 \item [INIT: An initiation to {\mmalfa}.] The INIT demo shows a few
  basic functions of MMAlpha. The demo comprises the following steps:
  loading and showing a program, normalizing, showing domains etc...
  \index{INIT demo}

 \item [FIR: Design of a FIR filter.]  The FIR demo shows the typical
steps for the synthesis of a linear architecture for a Finite Impulse
Response filter.  The demo comprises the following steps: loading the
initial program, pipelining, scheduling, assign parameter values,
control signal generation, {\alphaz} generation, {\alphard} generation.\index{FIR demo, Finite impulse
response filter}

 \item [MATVECT: architecture for matrix vector
multiplication.] 
The MATVECT demo shows the typical steps for the
synthesis of a linear architecture for a Matrix Vector Multiplication.
The demo comprises the following steps: loading the initial program,
pipelining, scheduling, assign parameter values, control signal
generation,  {\alphaz} generation, {\alphard} generation.\index{MATVECT demo,
Matrix vector architecture}

% \item [VERIF: Verification of the Matrix vector architecture.]
%The VERIF demo shows the verification of a linear architecture for a
%Matrix Vector Multiplication. The demo shows the use of subsystems,
%the inlining of subsystems, the use of libraries, C code generation
%and execution.\index{VERIF demo}

 \item [SVD: analysis of a Singular Value Decomposition
algorithm.]  The SVD demo shows the analysis of a larger Alpha
program. It shows the use of subsystems, the nlining subsystems, the
CNF code generation, the C code generation and execution.

% \item [SVDUPDATE: another demo on Singular Value Decomposition.]
%The SVD demo shows the analysis of another version of the SVD algorithm.

 \item [BINMULT: Synthesis of a multiplier.]  This demonstration
shows the synthesis of a bit-serial multiplier, from an initial
specification structured into several parameterized subsystems.  The
demo script contains the following steps: loading and displaying a
multi-system program, conversion into C code for simulation,
pipelining of the diffusions in the initial program, flattenning all
the systems for synthesis, scheduling before space/time
transformation, space/time transformation, pipelining of the inputs
(so that all the data enters the first processor in bit-serial
manner), simulation of the resulting array.

\item [ESTIM: synthesis of motion estimation] a two dimensionnal example, with 
simulation, on a real image, derivation of {\alphaz} code (warning, long demo).
\end{description}