### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/install.tex

\chapter{Installing \mmalfa{}}

\label{appendixtechset}

\label{install}

\newcommand{\mmalpha}{{MMalpha}}

\input{install1.tex}

\section{Getting started}

You may use either {\mma} under {\tt emacs}, or directly on the Unix

system.

\begin{itemize}

\item under {\tt emacs}, type {\tt  esc-X Mathematica}.

\item type math (use the {\mma} textual interface).

\item type mathematica (use the {\mma} textual interface).

\end{itemize}

%Do not use a Notebook, as the output of the pretty printer would not

%be available.

When the {\mma} session is started, you can type {\tt goDemo["INIT"]}

for the initiation demonstration.

When the {\mma} session is started, you can type {\tt on[]} or

\texttt{start[]} for the initiation demonstration.